---
title: "Reproducible Research: Peer Assessment 1"
author: "rwong013"
output: 
  html_document:
    keep_md: true
---

#Activity Monitoring

## Loading and preprocessing the data

The data is loaded here into the **activity** data table

```{r}
library(dplyr)
setwd("C:/Users/rwong/Documents/DSS/ReproducibleResearch/RepData_PeerAssessment1")
activity <- read.csv("C:/Users/rwong/Documents/DSS/ReproducibleResearch/RepData_PeerAssessment1/activity.csv")
```

## What is mean total number of steps taken per day?

First, we must sum the total number of steps per day. Then we will take the average and median total steps per day.
```{r}
library(ggplot2)
activitybyday <- activity %>% group_by(date) %>% summarize_each(funs(sum(., na.rm=TRUE)), steps)
meansteps <- floor(mean(activitybyday$steps, na.rm=TRUE))
mediansteps <- floor(median(activitybyday$steps, na.rm=TRUE))
qplot(data=activitybyday, date, steps, geom = "histogram", stat = "identity")
```

The mean total number of steps taken each day is `r meansteps`. 
The median number of steps taken each day is `r mediansteps`.

## What is the average daily activity pattern?

Using the ggplot2 library, we will plot the average of steps taken at each 5-minute interval in the day.

```{r}
activitybyinterval <- activity %>% group_by(interval) %>% summarize_each(funs(mean(., na.rm = TRUE)), steps)
qplot(data = activitybyinterval, interval, steps, geom = "line")

maxinterval <- activitybyinterval[activitybyinterval$steps == max(activitybyinterval$steps),]
```

On average, the most active 5-minute interval is `r maxinterval$interval` with `r maxinterval$steps` steps.


## Imputing missing values

```{r}
missingsteps <- sum(is.na(activity$steps))
```

The intervals with missing data is `r missingsteps`. We will fill in that missing data by filling NA values with the overall average for that interval.

```{r}
meanactivitybyday <- activity %>% group_by(date) %>% summarize_each(funs(sum(., na.rm=TRUE)), steps)
activityfill <- activity

for(i in 1:nrow(activityfill)){
        if(is.na(activityfill$steps[i])){
                activityfill$steps[i] <- activitybyinterval$steps[activitybyinterval$interval == activityfill$interval[i]]
        }
}

```

The effect that filling this data in is as follows:
```{r}
activityfillbyday <- activityfill %>% group_by(date) %>% summarize_each(funs(sum), steps)
meanstepsfill <- floor(mean(activityfillbyday$steps))
medianstepsfill <- floor(median(activityfillbyday$steps))
qplot(data=activityfillbyday, date, steps, geom = "histogram", stat = "identity")
```

Using the filled in data, the mean total number of steps taken each day is `r meanstepsfill` whereas previously it was `r meansteps`.
Using the filled in data, the median number of steps taken each day is `r medianstepsfill` whereas previously it was `r mediansteps`.

## Are there differences in activity patterns between weekdays and weekends?

We will add a new column that tells us which day of the week it is. This is done by checking if the date is a Saturday or Sunday, and labeling it as a weekend if it is, or a weekday if it is not.
```{r}
checkdate <- function(date){
        if(weekdays(as.Date(date)) == "Sunday" | weekdays(as.Date(date)) == "Saturday"){
                "Weekend"
        } else {
                "Weekday"
        }
}

activityfillday <- activityfill %>% rowwise() %>% mutate(weekday = checkdate(date))
```

Now we will plot the information and separate the information by weekday or weekend.
```{r}
activitybyweekday <- activityfillday %>% group_by(interval,weekday) %>% summarize_each(funs(mean(., na.rm = TRUE)), steps)
qplot(data = activitybyweekday, interval, steps, geom = "line", facets = weekday ~ .)
```